---
title: "Section 4: Amnion Markers"
date: "`r Sys.Date()`"
author:
  - name: Arun Seetharam
  - name: Ha Vu
    affiliation: Tuteja Lab
    affiliation_url: https://www.tutejalab.org
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: false
    lightbox: true
    gallery: true
    highlight: tango
---

```{r setup, include=FALSE}
options(max.print = "75")
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "assets/",
  fig.width = 8,
  prompt = FALSE,
  tidy = FALSE,
  message = FALSE,
  warning = TRUE
)
knitr::opts_knit$set(width = 75)
```
## Prerequisites

R packages required for this section are loaded.

```{r, warnings=TRUE, message=FALSE}
setwd("~/github/BAPvsTrophoblast_Amnion")
library(sva)
library(tidyverse)
library(DESeq2)
library(vsn)
library(pheatmap)
library(ggrepel)
library(RColorBrewer)
library(reshape2)
library(scales)
require(biomaRt)
library(EnhancedVolcano)
library(TissueEnrich)
library(corrplot)
library(rstatix)
library(ggpubr)
library(car)
library(FSA)
library(PupillometryR)
load("assets/myCorrplot.Rdata")
```

## Finding Amnion markers using `TissueEnrich`

```{r markers, warnings=TRUE, message=FALSE}
# Roost and Suzuki data used as different amnion tissues
cts.full <-
  as.matrix(read.csv(
    "assets/counts-batch-corrected-v6_tpm.txt",
    sep = "\t",
    row.names = "gene"
  ))
md.full <-
  read.csv(
    "assets/coldata-batch-corrected-v6.txt",
    sep = "\t",
    header = TRUE,
    row.names = 1
  )
mycts.t <- as.data.frame(t(cts.full))
metadata <- md.full[2]
merged.table <-
  merge(metadata, mycts.t, by = "row.names", all.x = TRUE)
merged.table <-
  merged.table %>% remove_rownames %>% column_to_rownames(var = "Row.names")
mycts.mean <- aggregate(. ~ group, merged.table, mean)
ordered <-
  as.data.frame(t(
    mycts.mean %>% remove_rownames %>% column_to_rownames(var = "group")
  ))
se <-
  SummarizedExperiment(
    assays = SimpleList(as.matrix(ordered)),
    rowData = row.names(ordered),
    colData = colnames(ordered)
  )


# results for TE enrichment, max tissues =2; fc threshold = 10; and expression threshold 1
output <- teGeneRetrieval(
  se,
  foldChangeThreshold = 10,
  maxNumberOfTissues = 2,
  expressedGeneThreshold = 1
)

# save
te.results.merged <- as.data.frame(assay(output))
write.table(
  te.results.merged,
  file = "enrichment-results_fc10_merged.tsv",
  sep = "\t",
  row.names = FALSE,
  quote = FALSE
)

full.ec.results <- as.data.frame(assay(output))
amnion.roost <- subset(full.ec.results, Tissue=='amnion' & Group=='Tissue-Enriched')
```


## Processing data for plotting heatmaps

This section uses the count data (selected datasets) generated in Section 1 for generating heatmaps. Briefly, the count data are imported in R, batch corrected using `ComBat_seq`, `vst` transformation is performed and normalized using `DESeq2`, and heatmaps are generated for selected genes.


### Dataset import
The `counts` data and its associated metadata (`coldata`) are imported for analyses.

```{r dataset, warnings=TRUE, message=FALSE}
counts = 'assets/counts-pca-v2.txt'
groupFile = 'assets/batch-pca-v2.txt'
coldata <-
  read.csv(
    groupFile,
    row.names = 1,
    sep = "\t",
    stringsAsFactors = TRUE
  )
cts <-
  as.matrix(read.csv(counts, sep = "\t", row.names = "gene.ids"))
```
Inspect the `coldata`.

```{r coldata}
DT::datatable(coldata)
```

Reorder columns of `cts` according to `coldata` rows. Check if the samples in both files match.

```{r order, warnings=TRUE, message=FALSE}
colnames(cts)
all(rownames(coldata) %in% colnames(cts))
cts <- cts[, rownames(coldata)]
```

### Batch correction

Using `ComBat_seq` (SVA package) to run batch correction - using bioproject IDs as variable (dataset origin).

```{r batchcorrect, warnings=TRUE, message=TRUE}
cov1 <- as.factor(coldata$authors)
adjusted_counts <- ComBat_seq(cts, batch = cov1, group = NULL)
all(rownames(coldata) %in% colnames(cts))
cts <- cts[, rownames(coldata)]
```

### Run DESeq2

The batch corrected read counts are then used for running DESeq2 analyses.

```{r deseq2, warnings=TRUE, message=FALSE}
dds <- DESeqDataSetFromMatrix(countData = adjusted_counts,
                              colData = coldata,
                              design = ~ group)
vsd <- vst(dds, blind = FALSE)
keep <- rowSums(counts(dds)) >= 10
dds <- dds[keep,]
dds <- DESeq(dds)
```

### Normalize counts

```{r normtable, warnings=TRUE, message=FALSE}
normalized_counts <- counts(dds, normalized = T) %>%
  data.frame() %>%
  rownames_to_column(var = "gene")
normalized_counts <- normalized_counts %>%
  as_tibble()
```

## Ensembl ID table

The gene lists have Ensembl gene-ID-version. We need them as gene-IDs. We also need other metadata later for these lists.
From Ensembl we will download metadata and attach to these lists.

```{r annotations, warnings=TRUE, message=FALSE}
ensembl = useMart("ENSEMBL_MART_ENSEMBL")
listDatasets(ensembl) %>%
  filter(str_detect(description, "Human"))
ensembl = useDataset("hsapiens_gene_ensembl", mart = ensembl)
listFilters(ensembl) %>%
  filter(str_detect(name, "ensembl"))
filterType <- "ensembl_gene_id_version"
filterValues <- rownames(cts)
attributeNames <- c('ensembl_gene_id_version',
                    'ensembl_gene_id',
                    'external_gene_name')
annot <- getBM(
  attributes = attributeNames,
  filters = filterType,
  values = filterValues,
  mart = ensembl
)
isDup <- duplicated(annot$ensembl_gene_id)
dup <- annot$ensembl_gene_id[isDup]
annot <- annot[!annot$ensembl_gene_id %in% dup, ]
```


## Required Functions

convert geneids function
```{r idsFun, warnings=TRUE, message=FALSE}
addGeneIds <- function(df) {
  merge(
    df,
    annot,
    by.x = "Gene",
    by.y = "ensembl_gene_id",
    all.x = TRUE,
    all.y = FALSE
  ) %>% drop_na() %>% filter(external_gene_name != "")
}
```
correlation function
```{r corFun, warnings=TRUE, message=FALSE}
cor.mtest <- function(mat, ...) {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat<- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], ...)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}
```
Data summary for the violin plots
```{r StatFun, warnings=TRUE, message=FALSE}
data_summary <- function(x) {
  m <- mean(x)
  ymin <- m-sd(x)
  ymax <- m+sd(x)
  return(c(y=m,ymin=ymin,ymax=ymax))
}
```
Theme for the plots
```{r themeFun, warnings=TRUE, message=FALSE}
# SOURCE: https://ourcodingclub.github.io/tutorials/dataviz-beautification/
theme_niwot <- function() {
  theme_bw() +
    theme(
      axis.text = element_text(size = 16),
      axis.title = element_text(size = 18),
      axis.line.x = element_line(color = "black"),
      axis.line.y = element_line(color = "black"),
      panel.border = element_blank(),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.grid.major.y = element_blank(),
      plot.margin = unit(c(1, 1, 1, 1), units = , "cm"),
      plot.title = element_text(
        size = 18,
        vjust = 1,
        hjust = 0
      ),
      legend.text = element_text(size = 12),
      legend.title = element_blank(),
      legend.position = c(0.95, 0.15),
      legend.key = element_blank(),
      legend.background = element_rect(
        color = "black",
        fill = "transparent",
        size = 2,
        linetype = "blank"
      )
    )
}
```
```{r filter1, warnings=TRUE, message=FALSE}
name1 <-
  c(
    "amnion_9w.1",
    "amnion_9w.2",
    "amnion_16w.2",
    "amnion_18w.1",
    "amnion_18w.2",
    "amnion_22w.1",
    "amnion_22w.2",
    "CT_7wk.1",
    "CT_7wk.2",
    "CT_9wk.1",
    "CT_11wk.1",
    "pBAP_D3.1",
    "pBAP_D3.2",
    "hESC_H9_BMP4_72h.1",
    "hESC_H9_BMP4_72h.2",
    "H1_gt70_D8_BAP.1",
    "H1_gt70_D8_BAP.2",
    "H1_gt70_D8_BAP.3"
  )
filter1 <- as.data.frame(name1)
name2 <-
  c(
    "BT_EVT_Okae.1",
    "BT_SCT_Okae.1",
    "BT_TSC_Okae.1",
    "BT_EVT_Okae.2",
    "BT_SCT_Okae.2",
    "BT_TSC_Okae.2",
    "CT_EVT_Okae.1",
    "CT_SCT_Okae.1",
    "CT_TSC_Okae.1",
    "CT_EVT_Okae.2",
    "CT_SCT_Okae.2",
    "CT_TSC_Okae.2",
    "CT_EVT_Okae.3",
    "CT_SCT_Okae.3",
    "CT_TSC_Okae.3",
    "amnion_18w.1",
    "amnion_9w.1",
    "amnion_18w.2",
    "amnion_16w.2",
    "amnion_22w.1",
    "amnion_9w.2",
    "amnion_22w.2"
  )
filter2 <- as.data.frame(name2)
```

```{r tableFun1, warnings=TRUE, message=FALSE}
getTable.main <- function(genelist = df,
                          data.to.plot = name1) {
  norm.filtered.table <-
    normalized_counts %>% filter(gene %in% genelist$ensembl_gene_id_version)
  norm.filtered.long <- melt(norm.filtered.table, id.vars = "gene")
  filter1 <- as.data.frame(data.to.plot)
  de_names <- filter1$data.to.plot
  norm.filtered.subset.long <-
    norm.filtered.long %>% filter (variable %in% de_names)
  colnames(norm.filtered.subset.long) <-
    c("gene", "condition", "norm.expression")
  norm.filtered.subset.long <-
    merge(
      norm.filtered.subset.long,
      annot,
      by.x = "gene",
      by.y = "ensembl_gene_id_version",
      all.x = TRUE,
      all.y = FALSE
    ) %>% drop_na() %>% filter(external_gene_name != "")
  norm.filtered.subset.long
}
```


```{r htmpFun, warnings=TRUE, message=FALSE}
runheatmap <-
  function(get_table = table,
           filename = "test_name",
           num = 20,
           data.to.plot = name1,
           ChartTitle = "Title") {
    norm.filtered.subset.long <- get_table
    norm.filtered.subset.table <-
      dcast(norm.filtered.subset.long,
            external_gene_name ~ condition,
            value.var = "norm.expression")
    write_delim(
      norm.filtered.subset.table,
      file = paste0(filename, "_expValues.tsv"),
      delim = "\t"
    )
    norm.filtered.subset.table <-
      norm.filtered.subset.table %>% column_to_rownames("external_gene_name")
    annotation <- data.frame(Condition = colnames(norm.filtered.subset.table))
    heat_colors <- brewer.pal(9, "YlOrRd")
    hmap.data <-
      as.matrix(norm.filtered.subset.table[, data.to.plot])
    g <- pheatmap(
      hmap.data,
      color = heat_colors,
      main = ChartTitle,
      cluster_rows = F,
      cluster_cols  = F,
      show_rownames = T,
      border_color = NA,
      fontsize = 14,
      scale = "row",
      fontsize_row = 10
    )
    g
  }
```




```{r corplotFun, warnings=TRUE, message=FALSE}
corPlot <- function(get_table = table,
                    ChartTitle = "Title",
                    mycolor = "darkorchid") {
  norm.filtered.subset.long <- get_table
  norm.filtered.subset.table <-
    dcast(norm.filtered.subset.long,
          external_gene_name ~ condition,
          value.var = "norm.expression")
  temp <- norm.filtered.subset.table[, -1]
  row.names(temp) <- norm.filtered.subset.table[, 1]
  norm.filtered.subset.table <- temp
  pcor <-
    cor(as.matrix(norm.filtered.subset.table), method = "spearman")
  p.mat <- cor.mtest(as.matrix(norm.filtered.subset.table))
  col <-
    colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
  my.corrplot(
    pcor,
    method = "color",
    col = col(200),
    type = "upper",
    order = "hclust",
    addCoef.col = mycolor,
    pch.cex = 1.5,
    p.mat = p.mat,
    insig = 'label_sig',
    sig.level = c(0.001, 0.01, 0.05),
    tl.col = "black",
    tl.srt = 45,
    number.cex = 0.8,
    tl.cex = 1,
    pch.col = "tomato",
    diag = FALSE,
    font.main = 4,
    mar = c(0, 0, 1, 0)
  )
}
```


```{r normFun, warnings=TRUE, message=FALSE}
normTestEach <- function(get_table = table,
                         ChartTitle = "Title") {
  norm.filtered.subset.long <- get_table
  ggplot(norm.filtered.subset.long, aes(x = norm.expression)) +
    stat_density(color = "darkblue", fill = "lightblue") +
    theme_minimal() +
    theme(
      axis.text.x = element_text(
        angle = 45,
        vjust = 1,
        hjust = 1,
        size = 12
      ),
      axis.text.y = element_text(size = 12),
      plot.margin = margin(10, 10, 10, 100),
      legend.position = "none",
      plot.title = element_text(
        color = "black",
        size = 18,
        face = "bold.italic"
      ),
      axis.line.x = element_line(
        colour = 'black',
        size = 0.5,
        linetype = 'solid'
      ),
      axis.ticks.x = element_line(
        colour = 'black',
        size = 1,
        linetype = 'solid'
      ),
      axis.title.x = element_text(
        color = "black",
        size = 14,
        face = "bold"
      ),
      axis.line.y = element_line(
        colour = 'black',
        size = 0.5,
        linetype = 'solid'
      ),
      axis.ticks.y = element_line(
        colour = 'black',
        size = 1,
        linetype = 'solid'
      ),
      axis.title.y = element_text(
        color = "black",
        size = 14,
        face = "bold"
      )
    )  +
    scale_y_continuous(expand = expansion(mult = c(0, .1)), breaks = pretty_breaks()) +
    facet_wrap( ~ condition, ncol = 5) +
    scale_x_log10()
}
```

```{r normAllFun, warnings=TRUE, message=FALSE}
normTestAll <- function(get_table = table,
                        ChartTitle = "Title") {
  norm.filtered.subset.long <- get_table
  ggplot(norm.filtered.subset.long, aes(x = norm.expression)) +
    stat_density(color = "purple", fill = "dodgerblue") +
    theme_minimal() +
    theme(
      axis.text.x = element_text(
        angle = 45,
        vjust = 1,
        hjust = 1,
        size = 12
      ),
      axis.text.y = element_text(size = 12),
      plot.margin = margin(10, 10, 10, 100),
      legend.position = "none",
      plot.title = element_text(
        color = "black",
        size = 18,
        face = "bold.italic"
      ),
      axis.line.x = element_line(
        colour = 'black',
        size = 0.5,
        linetype = 'solid'
      ),
      axis.line.y = element_line(
        colour = 'black',
        size = 0.5,
        linetype = 'solid'
      ),
      axis.ticks.x = element_line(
        colour = 'black',
        size = 1,
        linetype = 'solid'
      ),
      axis.ticks.y = element_line(
        colour = 'black',
        size = 1,
        linetype = 'solid'
      ),
      axis.title.x = element_text(
        color = "black",
        size = 14,
        face = "bold"
      ),
      axis.title.y = element_text(
        color = "black",
        size = 14,
        face = "bold"
      )
    )  +
    scale_y_continuous(expand = expansion(mult = c(0, .1)), breaks = pretty_breaks()) +
    scale_x_log10()
}
```
```{r addSampleFun, warnings=TRUE, message=FALSE}
addSampleInfo <- function(get_table = table) {
  norm.filtered.subset.long <- get_table
  metadata <- coldata %>%
    rownames_to_column(var = "condition") %>%
    select(condition, group)
  metadata$group <- sub("^(amnion_).*w", "Amnion", metadata$group)
  metadata$group <-
    sub("^(CT_7|CT_9|CT_11)wk", "CT", metadata$group)
  norm.filtered.subset.long <- merge(
    norm.filtered.subset.long,
    metadata,
    by.x = "condition",
    by.y = "condition",
    all.x = TRUE,
    all.y = FALSE
  )
  norm.filtered.subset.long
}
```

```{r vlnPlotFun, warnings=TRUE, message=FALSE}
vlnPlot <- function(info_table = Infotable) {
  norm.filtered.subset.long <- info_table
  ggplot(data = norm.filtered.subset.long,
         aes(x = group, y = norm.expression, fill = group)) +
    geom_flat_violin(position = position_nudge(x = 0.2, y = 0),
                     alpha = 0.8,
                     trim = FALSE) +
    geom_point(
      aes(y = norm.expression, color = group),
      position = position_jitter(width = 0.15),
      size = 1,
      alpha = 0.5
    ) +
    geom_boxplot(width = 0.2,
                 outlier.shape = NA,
                 alpha = 0.8) + stat_summary(
                   fun = mean,
                   geom = "point",
                   shape = 23,
                   size = 2
                 ) +
    labs(y = "\nNormalized Expression", x = NULL) +
    guides(fill = "none", color = "none") +
    scale_y_log10()  +
    theme_niwot() + theme(axis.text.x = element_text(
      angle = 45,
      vjust = 1,
      hjust = 1
    ))
}
```

## Prepare gene lists

Amnion markers genes using Roost dataset (generated using `TissueEnrich`)

```{r markers1, warnings=TRUE, message=FALSE}
amnion.roost.names <- addGeneIds(amnion.roost)
```

Amnion markers genes using Roost dataset (from Io _et al_., 2021)
```{r markers2, warnings=TRUE, message=FALSE}
file1 <-
  "~/TutejaLab/amnion_for_ms_20210715/heatmaps.v2/Fig6b_and_monkeygenes/genes-Fig6b.txt"
fig6b <-
  read.csv(file1,
           sep = "\t",
           header = TRUE,
           stringsAsFactors = FALSE)
fig6b <- merge(
  fig6b,
  annot,
  by.x = "genes",
  by.y = "external_gene_name",
  all.x = TRUE,
  all.y = FALSE
) %>% drop_na() 
```

## Format Conversion

```{r tableout1, warnings=TRUE, message=FALSE}
amnion.roost.new.main <- getTable.main(amnion.roost.names, name1)
amnion.roost.old.main <- getTable.main(fig6b, name1)
amnion.roost.new.supp <- getTable.main(amnion.roost.names, name2)
amnion.roost.old.supp <- getTable.main(fig6b, name2)
```

```{r tableout2, warnings=TRUE, message=FALSE}
amnion.roost.new.main.info <- addSampleInfo(amnion.roost.new.main)
amnion.roost.old.main.info <- addSampleInfo(amnion.roost.old.main)
amnion.roost.new.supp.info <- addSampleInfo(amnion.roost.new.supp)
amnion.roost.old.supp.info <- addSampleInfo(amnion.roost.old.supp)
```


## Visualization (new Amnion marker genes)


```{r hm1a, fig.cap="Fig 4.1A: Heatmap for the new amnion marker genes", fig.width=8, fig.height=20}
runheatmap(
  amnion.roost.new.main.info,
  data.to.plot = name1,
  num = 38,
  ChartTitle = "New amnion markers - expression across samples"
)
```

```{r cor1a, fig.cap="Fig 4.1B: Corelation plot for the new amnion marker genes", fig.width=12, fig.height=12}
corPlot(amnion.roost.new.main.info)
```
```{r normeach1a, fig.cap="Fig 4.1C: distribution plot for the new amnion marker genes", fig.width=8, fig.height=8}
normTestEach(amnion.roost.new.main.info)
```
```{r normall1a, fig.cap="Fig 4.1D: distribution plot for the new amnion marker genes", fig.width=8, fig.height=8}
normTestAll(amnion.roost.new.main.info)
```
```{r dt1a, warnings=TRUE, message=FALSE}
shapiro.each <-
  amnion.roost.new.main.info %>% dplyr::group_by(condition) %>% shapiro_test(norm.expression)
DT::datatable(shapiro.each,)
```
```{r dt2a, warnings=TRUE, message=TRUE}
shapiro.test(amnion.roost.new.main.info$norm.expression[0:5000])
```
```{r dt3a, warnings=TRUE, message=FALSE}
kruskal.test(norm.expression ~ group, data = amnion.roost.new.main.info)
```
```{r dt4a, warnings=TRUE, message=FALSE}
dunnTest <-
  dunnTest(norm.expression ~ group, data = amnion.roost.new.main.info, method =
             "bh")
g <- as.data.frame(dunnTest$res)
DT::datatable(g)
```
```{r vln1a, fig.cap="Fig 4.1E: Violin plot for the new amnion marker genes", fig.width=8, fig.height=8}
vlnPlot(amnion.roost.new.main.info)
```
  

## Visualization (old Amnion marker genes)


```{r hm1b, fig.cap="Fig 4.2A: Heatmap for the Io amnion marker genes", fig.width=8, fig.height=20}
runheatmap(
  amnion.roost.old.main.info,
  data.to.plot = name1,
  num = 20,
  ChartTitle = "Io amnion markers - expression across samples"
)
```

```{r cor1b, fig.cap="Fig 4.2B: Corelation plot for the Io amnion marker genes", fig.width=12, fig.height=12}
corPlot(amnion.roost.old.main.info, mycolor = "yellow")
```
```{r normeach1b, fig.cap="Fig 4.2C: distribution plot for the Io amnion marker genes", fig.width=8, fig.height=8}
normTestEach(amnion.roost.old.main.info)
```
```{r normall1b, fig.cap="Fig 4.2D: distribution plot for the Io amnion marker genes", fig.width=8, fig.height=8}
normTestAll(amnion.roost.old.main.info)
```
```{r dt1b, warnings=TRUE, message=FALSE}
shapiro.each <-
  amnion.roost.old.main.info %>% dplyr::group_by(condition) %>% shapiro_test(norm.expression)
DT::datatable(shapiro.each,)
```
```{r dt2b, warnings=TRUE, message=TRUE}
shapiro.test(amnion.roost.old.main.info$norm.expression[0:5000])
```
```{r dt3b, warnings=TRUE, message=FALSE}
kruskal.test(norm.expression ~ group, data = amnion.roost.old.main.info)
```
```{r dt4b, warnings=TRUE, message=FALSE}
dunnTest <-
  dunnTest(norm.expression ~ group, data = amnion.roost.old.main.info, method =
             "bh")
g <- as.data.frame(dunnTest$res)
DT::datatable(g)
```
```{r vln1b, fig.cap="Fig 4.2E: Violin plot for the Io amnion marker genes", fig.width=8, fig.height=8}
vlnPlot(amnion.roost.old.main.info)
```
  




## Visualization (new Amnion marker genes) for amnion and organoid datasets


```{r hm1c, fig.cap="Fig 4.3A: Heatmap for the new amnion marker genes", fig.width=8, fig.height=20}
runheatmap(
  amnion.roost.new.supp.info,
  data.to.plot = name2,
  num = 20,
  ChartTitle = "Io amnion markers - expression across samples"
)
```

```{r cor1c, fig.cap="Fig 4.3B: Corelation plot for the new amnion marker genes", fig.width=12, fig.height=12}
corPlot(amnion.roost.new.supp.info)
```
```{r normeach1c, fig.cap="Fig 4.3C: distribution plot for the new amnion marker genes", fig.width=8, fig.height=8}
normTestEach(amnion.roost.new.supp.info)
```
```{r normall1c, fig.cap="Fig 4.3D: distribution plot for the new amnion marker genes", fig.width=8, fig.height=8}
normTestAll(amnion.roost.new.supp.info)
```
```{r dt1c, warnings=TRUE, message=FALSE}
shapiro.each <-
  amnion.roost.new.supp.info %>% dplyr::group_by(condition) %>% shapiro_test(norm.expression)
DT::datatable(shapiro.each,)
```
```{r dt2c, warnings=TRUE, message=TRUE}
shapiro.test(amnion.roost.new.supp.info$norm.expression[0:5000])
```
```{r dt3c, warnings=TRUE, message=FALSE}
kruskal.test(norm.expression ~ group, data = amnion.roost.new.supp.info)
```
```{r dt4c, warnings=TRUE, message=FALSE}
dunnTest <-
  dunnTest(norm.expression ~ group, data = amnion.roost.new.supp.info, method =
             "bh")
g <- as.data.frame(dunnTest$res)
DT::datatable(g)
```
```{r vln1c, fig.cap="Fig 4.3E: Violin plot for the new amnion marker genes", fig.width=8, fig.height=8}
vlnPlot(amnion.roost.new.supp.info)
```



## Visualization (old Amnion marker genes) for amnion and organoid datasets


```{r hm1d, fig.cap="Fig 4.4A: Heatmap for the Io amnion marker genes", fig.width=8, fig.height=20}
runheatmap(
  amnion.roost.old.supp.info,
  data.to.plot = name2,
  num = 20,
  ChartTitle = "Io amnion markers - expression across samples"
)
```

```{r cor1d, fig.cap="Fig 4.4B: Corelation plot for the Io amnion marker genes", fig.width=12, fig.height=12}
corPlot(amnion.roost.old.supp.info, mycolor = "yellow")
```
```{r normeach1d, fig.cap="Fig 4.4C: distribution plot for the Io amnion marker genes", fig.width=8, fig.height=8}
normTestEach(amnion.roost.old.supp.info)
```
```{r normall1d, fig.cap="Fig 4.4D: distribution plot for the Io amnion marker genes", fig.width=8, fig.height=8}
normTestAll(amnion.roost.old.supp.info)
```
```{r dt1d, warnings=TRUE, message=FALSE}
shapiro.each <-
  amnion.roost.old.supp.info %>% dplyr::group_by(condition) %>% shapiro_test(norm.expression)
DT::datatable(shapiro.each,)
```
```{r dt2d, warnings=TRUE, message=TRUE}
shapiro.test(amnion.roost.old.supp.info$norm.expression[0:5000])
```
```{r dt3d, warnings=TRUE, message=FALSE}
kruskal.test(norm.expression ~ group, data = amnion.roost.old.supp.info)
```
```{r dt4d, warnings=TRUE, message=FALSE}
dunnTest <-
  dunnTest(norm.expression ~ group, data = amnion.roost.old.supp.info, method =
             "bh")
g <- as.data.frame(dunnTest$res)
DT::datatable(g)
```
```{r vln1d, fig.cap="Fig 4.4E: Violin plot for the Io amnion marker genes", fig.width=8, fig.height=8}
vlnPlot(amnion.roost.old.supp.info)
```

## Session Information

```{r sessioninfo}
sessionInfo()
```